### 2018-05-09 2_wmm.tex
Some additions (non-determinism, ...).

% TODO: the event-based representation is a non-deterministic model
%- The semantics of the event-flow graph is given:  "alglave: Don’t sit on the fence A static analysis approach to automatic fence insertion" page 8 or 9
%"To build the aeg , we define a semantics of goto-programs in terms of abstract events, ..."
% TODO: OR from here: http://www.cprover.org/etaps/esop.pdf "Software Verification for Weak Memory via Program Transformation"

>>>>

There are three main types of sources of non-determinism in concurrent programs~\cite{musuvathi2008fair}:
\begin{enumerate}[noitemsep,topsep=0pt]
\item \textit{input non-determinism}, which is a standard undecidable problem for all static analysis methods: to resolve the user input, system call from the environment, unresolved function calls, etc.;
\item \textit{scheduling non-determinism}, caused by the interleavings, which in turn are caused by the scheduler activity;
\item \textit{memory-model non-determinism}, caused by hardware and compiler relaxations.
\end{enumerate}

The event-based program model is able to emulate effectively the second and the third types of sources of non-determinism, while the first one can be coped by standard static analysis methods~\cite{landi1992undecidability,SurveySymExec-CSUR18}.

%- The semantics of the event-flow graph is given:  "alglave: Don’t sit on the fence A static analysis approach to automatic fence insertion" page 8 or 9
%"To build the aeg , we define a semantics of goto-programs in terms of abstract events, ..."
% TODO: OR from here: http://www.cprover.org/etaps/esop.pdf "Software Verification for Weak Memory via Program Transformation"

### 2018-05-09 4_impl.tex
Some additions (non-determinism, ...).

As its predecessor, \porthos[2] uses the open-source SMT-solver \tool{Z3}%
\footnote{The \tool{Z3} project repository: \url{https://github.com/Z3Prover/z3}}%
by Microsoft Research~\cite{de2008z3}. However, unlikely its predecessor, the \porthos[2] has an additional abstraction level \zformula{} (see Section~\ref{ch:impl:model:zformula}) that allow to use any other SMT-solver. %TODO: *should* allow?

@@

All \ytree{} expressions implement the \texttt{YExpression} interface.
On the \ytree{} level, the pointer arithmetic is modelled by the integer number \textit{pointer level} of an expression (although in fact this is the property of a type not of an expression, the y-tree is an untyped syntax tree, therefore the elements \ytree{} should carry this property).
We distinguish the subset of expressions that imply no side-effects, they implement the interface \texttt{YAtom} and can be global or local (which is defined also syntactically).
%\textbf{(TODO:THIS SHOULD BE IN PRE-PROCESSING)} %TODO <--

- carries the origin (text) -> all the origins can be resolved by the \texttt{LocationService} %todo: rename from CodeCitationService + create 3 instances for each of level. Or keep it for Y->text, and store references to the original Y-object by X-objects

@@

The \ytree{} is compiled into the low-level event-based program representation called \textit{\xgraph{}}.
The mathematical structure of event-flow graph was discussed in Section~\ref{ch:wmm:event}.
%In general, \xgraph{} follows the structure described there, however it is constructed in several stages.
The nodes of the graph are events, and the edges are basic relations: the control-flow relation \po and the data-flow relations \co and \rf.
Hereinafter, we denote the \xgraph{} with only control-flow edges as \xgraph[CF], the \xgraph{} with only data-flow edges as \xgraph[CF]. The complete \xgraph{} is $\xgraph[CF+DF] = \xgraph[CF] \; \cup \; \xgraph[DF]$.
\ Figure~\ref{fig:class-diagrams:XEntity-interfaces} represents the main hierarchy of the X-abstraction level.

@@

An event (\texttt{XEvent}) represents the fact of executing the primitive operation, which is independent from other events.
Each event is specified by the process generated them and a unique event label.
This information is stored by events in the immutable structure \texttt{XEventInfo}.

- carries the origin (Y)

@@

  \1 \texttt{XFenceEvent} \\
    The fences are implemented as an enumeration \texttt{XBarrierEvent}.
    Current implementation of \porthos[2] supports all fences supported by \porthos{}: 
    \texttt{mfence}, %TODO: what about lfence and sfence from the x86fences.cat ?
    \texttt{sync}, \texttt{optsync}, \texttt{lwsync}, \texttt{optlwsync}, \texttt{ish}, \texttt{isb}, and \texttt{isync}.%TODO: description of each

@@

\paragraph{Graph builder} % TODO: in code: remove graph hierarchy by combining unrolled and normal. Add a bool 'UnrolledStatus' -- thus we can keep genericity
\label{ch:impl:model:xgraph:builder}
%TODO: code : NOTE: 'FlowGraph<N extends FlowGraphNode>' may be just 'FlowGraph<N>'

\textbf{todo}

>>>>

As its predecessor, \porthos[2] uses the open-source SMT-solver \tool{Z3} %
\footnote{The \tool{Z3} project repository: \url{https://github.com/Z3Prover/z3}}%
by Microsoft Research~\cite{de2008z3}. However, unlikely its predecessor, the \porthos[2] has an additional abstraction level \zformula{} (see Section~\ref{ch:impl:model:zformula}) that allow to use any other SMT-solver. %TODO: *should* allow?

@@

All \ytree{} expressions implement the \texttt{YExpression} interface.
On the \ytree{} level, the pointer arithmetic is modelled by the integer number \textit{pointer level} of an expression (although in fact this is the property of a type not of an expression, the y-tree is an untyped syntax tree, therefore the elements \ytree{} should carry this property).
We distinguish the subset of expressions that imply no side-effects, they implement the interface \texttt{YAtom} and can be global or local (which is defined also syntactically).
Each \ytree{} expression carries the original code coordinates, that can be resolved to text by the service \texttt{LocationService}. %todo: rename from CodeCitationService + create 3 instances for each of level. Or keep it for Y->text, and store references to the original Y-object by X-objects

%\textbf{(TODO:THIS SHOULD BE IN PRE-PROCESSING)} %TODO <--

@@

The \ytree{} is compiled into the low-level event-based program representation called \textit{\xgraph{}}.
The mathematical structure of event-flow graph was discussed in Section~\ref{ch:wmm:event}.
%In general, \xgraph{} follows the structure described there, however it is constructed in several stages.
The nodes of the graph are events, and the edges are basic relations: the control-flow relation \po and the data-flow relations \co and \rf.
Hereinafter, we denote the \xgraph{} with only control-flow edges as \xgraph[CF], the \xgraph{} with only data-flow edges as \xgraph[CF].
The complete \xgraph{} is $\xgraph[CF+DF] = \xgraph[CF] \; \cup \; \xgraph[DF]$.
The UML diagraph on Figure~\ref{fig:class-diagrams:XEntity-interfaces} represents the hierarchy of main interfaces of the X-abstraction level.
%The \xgraph{} is constructed by the mutable graph builder. %TODO: name of builder -- after refactoring

@@

An event (\texttt{XEvent}) represents the fact of executing the primitive operation, which is independent from other events.
Each event is specified by the process generated them and a unique event label.
This information is stored by events in the immutable structure \texttt{XEventInfo}.
Also, each event carries the reference to the Y-instruction that has generated it.

@@

  \1 \texttt{XFenceEvent} \\
    The fences are implemented as an enumeration \texttt{XBarrierEvent}.
    Current implementation of \porthos[2] supports all fences supported by \porthos{}: 
    \texttt{mfence}, %TODO: what about lfence and sfence from the x86fences.cat ?
    \texttt{sync}, \texttt{optsync}, \texttt{lwsync}, \texttt{optlwsync}, \texttt{ish}, \texttt{isb}, and \texttt{isync}.%TODO: description of each!

@@

%\paragraph{Graph builder} % TODO: in code: remove graph hierarchy by combining unrolled and normal. Add a bool 'UnrolledStatus' -- thus we can keep genericity
%\label{ch:impl:model:xgraph:builder}
%TODO: code : NOTE: 'FlowGraph<N extends FlowGraphNode>' may be just 'FlowGraph<N>'
